---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, echo=TRUE, message=FALSE}

setwd("~/Documents/Data and Society/Data and Society")

library(tidyverse)
library(readr)
library(stringr)
library(geojsonR)
library(leaflet) 
library(RColorBrewer)
library(ggmap)
library(sf)
library(wesanderson)
library(corrr)
library(ggcorrplot)
library(FactoMineR)
library(factoextra)


job_access_gap_dict <- read_csv("job_access_gap_dict.csv")
job_access_gap_2_ <- read_csv("job_access_gap (2).csv")
job_access_gap_equity <- read_csv("job_access_gap_equity.csv")
job_access_gap_equity_dict <- read_csv("job_access_gap_equity_dict.csv")
job_access_geojson <- FROM_GeoJson(url_file_string = "https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2021/04/15/job_access_gap.geojson")
varblocks2018<- read_csv("varblocks2018.csv")
finality3 <- read_csv("finality3.csv")

access_json <- read_sf("https://urban-data-catalog.s3.amazonaws.com/drupal-root-live/2021/04/15/job_access_gap.geojson") %>%
  dplyr::filter(MSA == "Seattle" | MSA == "Baltimore") %>%
  mutate(GEOID = as.double(GEOID))

Spatial <- read_csv("Spatial.csv")
  


```

### Exercise 1



```{r code-chunk-label, echo=TRUE}

job_access_gap_new <- left_join(job_access_gap_equity, job_access_gap_2_, by= c('cbsa_name' = 'MSA'))

Access_new<-job_access_gap_new %>%
    pivot_longer(cols = c(pctasian, pctblack, pcthispanic,pctwhite, pctother, pctnonwhite), names_to = "race", values_to= "value")

job_access_gap_Seattle<- job_access_gap_new %>%
  filter(cbsa_name == "Seattle") %>%
  relocate(cbsa_name) %>%
  select(cbsa_name, Access30Transit, OverallJobAccess, pctasian, pctblack, pcthispanic,pctwhite,pctother,pctnonwhite, spatialmismatch)


boxplot(Access30Transit~cbsa_name, data = job_access_gap_new)


boxplot(OverallJobAccess~cbsa_name, data = job_access_gap_new)

boxplot(spatialmismatch ~ cbsa_name , data=job_access_gap_new)


```


```{r eval=FALSE, include=FALSE}

race_groupedSeattle<-job_access_gap_equity %>%
  select(type, pctasian,pctblack,pcthispanic,pctwhite,cbsa_name) %>%
  filter(cbsa_name == "Seattle" ) %>%
  pivot_longer(c(pctasian,pcthispanic, pctblack,pctwhite),
               names_to = "Race",
               values_to = "Percentages")

race_groupedBaltimore<-job_access_gap_equity %>%
  select(type, pctasian,pctblack,pcthispanic,pctwhite,cbsa_name) %>%
  filter(cbsa_name == "Baltimore" ) %>%
  pivot_longer(c(pctasian,pcthispanic, pctblack,pctwhite),
               names_to = "Race",
               values_to = "Percentages")
  
  

  
  
  race_groupedSeattle%>%
  ggplot(mapping = aes(x=Race, y= Percentages , fill = type)) +
  geom_bar( position = "dodge", stat = "identity", ) +
  labs(title = "Comparing Spatial Mismatch in Seattle",
       subtitle = "General versus Worst Grouped by Race",
       x = "Racial Groups",
       y= "Percentages")+
  scale_fill_manual(values=wes_palette(n=4, name="GrandBudapest2"))+
    scale_x_discrete( limits = c("pctwhite", "pctasian", "pcthispanic", "pctblack"),
                      labels = c("White", "Asian","Hispanic", "Black")
                    
    )

  
  race_groupedBaltimore%>%
  ggplot(mapping = aes(x=Race, y= Percentages , fill = type)) +
  geom_bar( position = "dodge", stat = "identity", ) +
  labs(title = "Comparing Spatial Mismatch in Baltimore",
       subtitle = "General versus Worst Grouped by Race",
       x = "Racial Groups",
       y= "Percentages")+
      scale_fill_manual(values=wes_palette(n=4, name="GrandBudapest2"))+
      scale_x_discrete( limits = c("pctwhite", "pctblack", "pcthispanic", "pctasian"),
                      labels = c("White", "Black","Hispanic", "Asian")
                    
    )
```


```{r eval=FALSE, include=FALSE}


POCwhite_equitySeattle<-job_access_gap_equity %>%
  select(type,pctwhite, pctnonwhite, cbsa_name) %>%
  filter(cbsa_name == "Seattle" ) %>%
  pivot_longer(c(pctwhite,pctnonwhite),
               names_to = "Race",
               values_to = "Percentages")


POCwhite_equitySeattle%>%
  ggplot(mapping = aes(x=Race, y= Percentages , fill = type)) +
  geom_bar( position = "dodge", stat = "identity", ) +
  labs(title = "Comparing Spatial Mismatch in Seattle",
       subtitle = "General versus Worst Grouped by Race",
       x = "Racial Groups",
       y= "Percentages") +
   scale_fill_manual(values=wes_palette(n=4, name="GrandBudapest2"))+
    scale_x_discrete( limits = c("pctwhite", "pctnonwhite"),
                      labels = c("White", "People of Color")
                    
    )




POCwhite_equityBaltimore<-job_access_gap_equity %>%
  select(type,pctwhite, pctnonwhite, cbsa_name) %>%
  filter(cbsa_name == "Baltimore" ) %>%
  pivot_longer(c(pctwhite,pctnonwhite),
               names_to = "Race",
               values_to = "Percentages")


POCwhite_equityBaltimore%>%
  ggplot(mapping = aes(x=Race, y= Percentages , fill = type)) +
  geom_bar( position = "dodge", stat = "identity", ) +
  labs(title = "Comparing Spatial Mismatch in Baltimore",
       subtitle = "General versus Worst Grouped by Race",
       x = "Racial Groups",
       y= "Percentages") +
   scale_fill_manual(values=wes_palette(n=4, name="GrandBudapest2"))+
    scale_x_discrete( limits = c("pctnonwhite", "pctwhite"),
                      labels = c("People of Color", "White")
    )
                    


```


### Exercise 2

```{r echo=TRUE}


finality<-finality3 %>%
 select(GEOID, spatialmismatch, MSA, geometry, WhiteProp, BlackProp, AsianProp, HispanicProp,hsdiplomaProp, hsnodiplomaProp, college1Prop, somecollegeProp, assProp, bachelorProp ) 


Seattle_finality<- finality %>%
  filter(MSA == "Seattle")

Baltimore_finality<- finality %>%
  filter(MSA == "Baltimore")

Spatial <- Spatial %>%
  select(MSA,hsdiplomaProp, college1Prop, somecollegeProp, assProp, bachelorProp, GEDProp, p12Prop, WhiteProp, POCProp) 

Seattle_spatial <- Spatial %>%
  filter(MSA == "Seattle") %>%
  select(-MSA)
  

Baltimore_spatial <- Spatial %>%
  filter(MSA == "Baltimore")%>%
  select(-MSA)
  






```





### Overall 


```{r eval=FALSE, include=FALSE}

cor.test(finality$spatialmismatch, finality$WhiteProp)

overallm1<-lm(spatialmismatch~WhiteProp, data= finality)

summary(overallm1)

```

### Seattle 

Observing how raceproportion significance on spatial mismatch with correlation testing. 

```{r eval=FALSE, include=FALSE}

cor.test(Seattle_finality$spatialmismatch, Seattle_finality$WhiteProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$BlackProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$AsianProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$HispanicProp)


Seattle_raceall<- lm(spatialmismatch~ WhiteProp +BlackProp +AsianProp +HispanicProp, data = Seattle_finality)
summary(Seattle_raceall)

Seattle_racesome <- lm(spatialmismatch~WhiteProp+AsianProp, data = Seattle_finality)
summary(Seattle_racesome)


```

```{r eval=FALSE, include=FALSE}

cor.test(Seattle_finality$spatialmismatch, Seattle_finality$hsdiplomaProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$hsnodiplomaProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$college1Prop)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$somecollegeProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$assProp)
cor.test(Seattle_finality$spatialmismatch, Seattle_finality$bachelorProp)


Seattle_edall<- lm(spatialmismatch~ hsdiplomaProp + hsnodiplomaProp + college1Prop + somecollegeProp + assProp+ bachelorProp, data = Seattle_finality)
summary(Seattle_edall)

Seattle_edsome<- lm(spatialmismatch~ hsdiplomaProp + college1Prop + bachelorProp, data = Seattle_finality)
summary(Seattle_edsome)

Seattle_edrefined<- lm(spatialmismatch ~ college1Prop +bachelorProp, data = Seattle_finality)




```

```{r eval=FALSE, include=FALSE}


Seattle_all <- lm(spatialmismatch ~ WhiteProp + BlackProp +AsianProp + HispanicProp + hsdiplomaProp + hsnodiplomaProp + college1Prop + somecollegeProp + assProp+ bachelorProp, data = Seattle_finality)

step.model <- MASS::stepAIC(Seattle_all, direction = "both", 
                      trace = FALSE)
summary(step.model)
```



### Baltimore

```{r eval=FALSE, include=FALSE}
cor.test(Baltimore_finality$spatialmismatch, Baltimore_finality$WhiteProp)
cor.test(Baltimore_finality$spatialmismatch, Baltimore_finality$BlackProp)
cor.test(Baltimore_finality$spatialmismatch, Baltimore_finality$AsianProp)
cor.test(Baltimore_finality$spatialmismatch, Baltimore_finality$HispanicProp)


Baltimore_raceall<- lm(spatialmismatch~ WhiteProp +BlackProp +AsianProp +HispanicProp, data = Baltimore_finality)
summary(Baltimore_raceall)

Baltimore_racesome <- lm(spatialmismatch~BlackProp + AsianProp + HispanicProp, data = Baltimore_finality)
summary(Baltimore_racesome)

Baltimore_racewb<- lm(spatialmismatch~ WhiteProp + BlackProp, data = Baltimore_finality)
summary(Baltimore_racewb)



```


```{r eval=FALSE, include=FALSE}
Baltimore_all <- lm(spatialmismatch ~ WhiteProp + BlackProp +AsianProp + HispanicProp + hsdiplomaProp + hsnodiplomaProp + college1Prop + somecollegeProp + assProp+ bachelorProp, data = Baltimore_finality)

Baltimore_allmodel <- MASS::stepAIC(Seattle_all, direction = "both", 
                      trace = FALSE)
summary(Baltimore_all)

```


```{r include=FALSE}
census_access <- left_join(access_json, finality3, by = "GEOID")


space <- colorBin(
  palette = "RdYlBu",
  domain = census_access$spatialmismatch.x)
```



```{r eval=FALSE, include=FALSE}

cutpoints = c(0, 0.05, 0.10, 0.15,0.20, 0.30, 0.50,0.70,1.00)

labels= c("0-0.05", "0.06-0.10", "0.11-0.15", "0.16-0.30", "0.31-0.50", "0.51-0.70","0.71-1.00")

my_colors <- c("#ffffd4", "#fee391", "#fec44f", "#fe9929", "#ec7014", "#cc4c02" , "#8c2d04")



degree <- colorBin(palette = my_colors, domain = census_access$bachelorProp, bins = cutpoints)

census_access %>%
  dplyr::filter(MSA.y == "Seattle") %>%
  sf::st_transform('+proj=longlat +datum=WGS84') %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons(color = "#444444", 
              weight = 0.5,
              smoothFactor = 0.5, 
              fillColor=~degree(round(census_access$bachelorProp,3)), 
              fillOpacity = 0.8, 
              highlightOptions = highlightOptions(color = "white", 
                                                  weight = 2,
                                                  bringToFront = TRUE), 
              label = round(census_access$bachelorProp, 3), 
              labelOptions = labelOptions(
                    style = list("font-weight" = "normal", padding = "3px 8px"),
                  textsize = "15px",
                direction = "auto")) %>%
  addLegend("bottomright",
            pal = degree, 
            values = census_access$bachelorprop,
            title = "BachelorProp",
             opacity = 1
  )


```

### PCA

## Seattle 





```{r include=FALSE}

Seattle_finality3 <- Seattle_finality %>%
  select(WhiteProp,AsianProp, HispanicProp, BlackProp, hsdiplomaProp,hsnodiplomaProp,somecollegeProp,college1Prop,assProp,bachelorProp)


#Normalizing Data 

normalized_Seattle<- scale(Seattle_finality3)

#Compute Correlation Matrix

Seattlecorr_matrix <- cor(normalized_Seattle)
ggcorrplot(Seattlecorr_matrix, colors = c("#d8b365", "#f5f5f5", "#5ab4ac"))

#Application 

PCA_Seattle <-princomp(Seattlecorr_matrix)
summary(PCA_Seattle)

fviz_pca_var(PCA_Seattle, col.var = "cos2", gradient.cols = c("black", "orange", "green"),repel = TRUE)

PCA_Seattle$loadings[, 1:2]

Seattlepcs<-as.matrix(normalized_Seattle %*% PCA_Seattle$loadings[, 1:2])
Seattlepc <- prcomp(Seattle_finality3, center = TRUE, scale. = TRUE)

Seattle2 <- as.data.frame(cbind(spatialmismatch=Seattle_finality$spatialmismatch, Seattlepc$x[,1:2]))



```

```{r include=FALSE}

#Seattle Linear Regression with PCA

M1_PCA_Seattle <- lm(spatialmismatch ~ PC1 + PC2, data = Seattle2)

M11_PCA_Seattle <-lm(spatialmismatch ~ PC1, data=Seattle2)


summary(M1_PCA_Seattle)
summary(M11_PCA_Seattle)

qqnorm(M1_PCA_Seattle$resid)
qqnorm(M11_PCA_Seattle$resid)

```

##Seattle Model 2
$$ \widehat{Spatial Mismatch} = e^{-0.122946 \cdot PC1  -2.663302} $$


This has an $$ R^{2}_{adj} = 0.16$$ We used a logarthmic transformation on 
spatial mismatch since the linearity model had skewness to the right. Additionally,
the normality probability plot suggests a slight upward trend reminsicient of a 
log/polynomial transformation. 

Furthermore,
our current $$ R^{2}_{adj} $$  has increased from $$ R^{2}_{adj} = 0.13$$. 
Despite the smaller increase our our predictors we have decieded to go with 
a simpler model because of the readibility in the interaction between Principal 
Component 1 and Spatial Mismatch. Recall that Principal Component 1 main contributors were our original variables $\textit{BachelorProp}$ and $\textit{hsdiplomaProp}$. These variables contrasted each other. The variable $\textit{hsdiplomaProp}$ has a negative weight of $-0.43$ which suggests that as the proportion of people who have a high school diploma than typical increase Principal Component 1 will decrease. Since Princpal Component 1 has a negative weight in our overall model, we can say that that there are a higher 
proportion than average of people who have a high school diploma. 


```{r echo=FALSE, warning=FALSE}

qqnorm(M11_PCA_Seattle$resid)
qqline(M11_PCA_Seattle$fitted)
 hist(resid(M11_PCA_Seattle))

plot(M11_PCA_Seattle$resid- M11_PCA_Seattle$fitted)


Seattle2$logspatialmismatch<-log(Seattle2$spatialmismatch)
Seattle2$logPC1 <- log(Seattle2$PC1)

Seattle2[is.na(Seattle2)| Seattle2 == "Inf"| Seattle2 == "-Inf"] = NA



M12_PCA_Seattle <-lm(logspatialmismatch ~ PC1, data=Seattle2)
summary(M12_PCA_Seattle)

qqnorm(M12_PCA_Seattle$resid)
hist(M12_PCA_Seattle$resid)
plot(M12_PCA_Seattle$resid ~ M12_PCA_Seattle$fitted)

M13_PCA_Seattle <- lm(logspatialmismatch ~logPC1, data=Seattle2)
summary(M13_PCA_Seattle) #Don't use this model, logPCI has less values. 

car::qqPlot(M12_PCA_Seattle)


```

Analysis Seattle Best Model: 

Just as a reminder Principal Component 1 for Seattle heaviest weights consist of the 
variables $\textit{hsdiplomaProp}$ which is the proportion of low-income
people in block groups that have attained a high school education and $\textit{bachelorProp}$ which is the 
proportion of low-income people in block groups that have attained a bachelors. 

We see the proportion of low-income people who have attained 
their high school diploma in block groups is higher than average, the higher the
spatial mismatch is. Especially looking at the graph when the Principal component 
One is between -2 and 0, spatial mismatch tends to increase. 

Then as we see the proportion of low-income people who have attained their bacelor's 
in block groups is higher than average, the lower the spatial mismatch is. We 
see that as get closer to 4, the spatial mismatch is close to 0. 

Something interesting to note, is that amongst principal component 1, is that the variable $\textit{HispanicrProp}$, $\textit{BlackProp}$ and $\textit{WhiteProp}$
are all relatively high as well. With $\textit{HispanicrProp} = -0.37 $, $\textit{BlackProp} =  -0.29$  and $\textit{WhiteProp} = 0.29$. In this case, 
we also see that as Hispanic proportions are higher than average in block groups, 
principal component 1 will decrease. Similarly, as Black proportions are higher 
than average in block groups principal component 1 will also decrease. Converslely
as white proportions are higher than average in block groups, princpal component 1
will also increase. 

We can say that in general we can say that principal component 1 primarily is seeing 
characteristics of education, with the heavy emphasis on high school diplomas and 
bachelor's degrees. 





```{r echo=FALSE}
plot(spatialmismatch ~ PC1, data=Seattle2,
     xlab = "Principal Component 1", 
     ylab = "Spatial Mismatch")
curve(exp(coef(M12_PCA_Seattle)[2]*x + coef(M12_PCA_Seattle)[1]), lwd=2, add =T, col ="blue")

#Use curve fitting, on normal variables using exponents. 

```

## Seattle 3

```{r eval=FALSE, include=FALSE}

Seattle2$sqrtspatialmismatch<-sqrt(Seattle2$spatialmismatch)
Seattle2$sqrtPC1 <- sqrt(Seattle2$PC1)

M14_PCA_Seattle <-lm(sqrtspatialmismatch ~ PC1, data=Seattle2)



qqnorm(M14_PCA_Seattle$resid)
hist(M14_PCA_Seattle$resid)

summary(M14_PCA_Seattle)

summary(lm(spatialmismatch ~ PC1 + I(PC1^2), data = Seattle2))


```


### Seattle Spatial PCA 

Exact same model as orginal, despite different transformations

```{r}

#Normalizing Data 

normalized_SeattleSpatial<- scale(Seattle_spatial)

#Compute Correlation Matrix

SeattleSpatialcorr_matrix <- cor(normalized_SeattleSpatial)
ggcorrplot(SeattleSpatialcorr_matrix, colors = c("#d8b365", "#f5f5f5", "#5ab4ac"))

#Application 

PCA_SeattleSpatial <-princomp(SeattleSpatialcorr_matrix)
summary(PCA_SeattleSpatial)

fviz_pca_var(PCA_SeattleSpatial, col.var = "black")

PCA_SeattleSpatial$loadings[, 1:2]


SeattleSpatialpcs<-as.matrix(normalized_SeattleSpatial %*% PCA_SeattleSpatial$loadings[, 1:2])
SeattleSpatialpc <- prcomp(Seattle_finality3, center = TRUE, scale. = TRUE)

Seattle3 <- as.data.frame(cbind(spatialmismatch=Seattle_finality$spatialmismatch, SeattleSpatialpc$x[,1:2]))



```


###Seattle Spatial Model 1


PC2 is not significant we are going to ignore it and focus on the relationship 
PC1 and spatial mismatch. 


```{r}

#Seattle Spatial Linear Regression with PCA

M2_PCA_SeattleSpatial <- lm(spatialmismatch ~ PC1 + PC2, data = Seattle3)

M21_PCA_SeattleSpatial <-lm(spatialmismatch ~ PC1, data=Seattle3)



summary(M21_PCA_SeattleSpatial)


```


Analyzing Residuals

```{r}
qqnorm(M21_PCA_SeattleSpatial$resid)
hist(M21_PCA_SeattleSpatial$resid)

#trasnformation 

Seattle3$logspatialmismatch<- log(Seattle3$spatialmismatch)

Seattle3[is.na(Seattle3)| Seattle3 == "Inf"| Seattle3 == "-Inf"] = NA

M22_PCA_SeattleSpatial<- lm(logspatialmismatch~PC1, data=Seattle3)
summary(M22_PCA_SeattleSpatial)


qqnorm(M22_PCA_SeattleSpatial$resid)






```

```{r}


summary(lm(spatialmismatch ~ PC1 + I(PC1^2), data = Seattle3))

```






```{r}


Baltimore_finality3 <- Baltimore_finality %>%
  select(WhiteProp,AsianProp, HispanicProp, BlackProp, hsdiplomaProp,hsnodiplomaProp,somecollegeProp,college1Prop,assProp,bachelorProp)


#Normalizing Data 

normalized_Baltimore<- scale(Baltimore_finality3)

#Compute Correlation Matrix

Baltimorecorr_matrix <- cor(normalized_Baltimore)
ggcorrplot(Baltimorecorr_matrix, colors = c("#d8b365", "#f5f5f5", "#5ab4ac"))

#Application 

PCA_Baltimore <-princomp(Baltimorecorr_matrix)
summary(PCA_Baltimore)

PCA_Baltimore$loadings[, 1:2]

fviz_pca_var(PCA_Baltimore, col.var = "cos2", gradient.cols = c("black", "orange", "green"),repel = TRUE)


Baltimorepcs<-as.matrix(normalized_Baltimore %*% PCA_Baltimore$loadings[, 1:2])

Baltimorepc <- prcomp(Baltimore_finality3, center = TRUE, scale. = TRUE)

Baltimore2 <- as.data.frame(cbind(spatialmismatch=Baltimore_finality$spatialmismatch, Baltimorepc$x[,1:2]))

 

```



```{r}
# Baltimore Fitting Linear Model PCA


M1_PCA_Baltimore <- lm(spatialmismatch ~ PC1 + PC2, data = Baltimore2)

M11_PCA_Baltimore <-lm(spatialmismatch ~ PC2, data=Baltimore2)


summary(M1_PCA_Baltimore)

summary(M11_PCA_Baltimore)

qqnorm(M11_PCA_Baltimore$resid)
hist(M11_PCA_Baltimore$resid)




```
### Baltimore Model 2 

Too much, skew to the left. 
```{r, warning=FALSE}

Baltimore2$logspatialmismatch<- log(Baltimore2$spatialmismatch)

Baltimore2[is.na(Baltimore2)| Baltimore2 == "Inf"| Baltimore2 == "-Inf"] = NA

M12_PCA_Baltimore<- lm(logspatialmismatch ~ PC2, data= Baltimore2)

summary(M12_PCA_Baltimore)
hist(M12_PCA_Baltimore$resid)

```

###Baltimore Model 3

$$ \widehat{Spatial Mismatch} = ({-0.0006 \cdot PC2 + 0.282})^2 $$

This has an $$ R^{2}_{adj} = 0.01$$ We used a square root  transformation on 
spatial mismatch since the linearity model had skewness to the right and 
logarthmic transfomration was too much. 

Furthermore,
our current $$ R^{2}_{adj} =0.02$$  has decreased to $$ R^{2}_{adj} = 0.01$$. 
but we are able to interpret this model with more simplicity. 
Despite the smaller increase our our predictors we have decieded to go with 
a simpler model because of the readability in the interaction between Principal 
Component 2 and Spatial Mismatch. Recall that Principal Component 2 main contributors 
were our original variables $\textit{BachelorProp}$ ,  $\textit{WhiteProp}$ ,and 
$\textit{BlackProp}$. These variables contrasted each other. The variable $\textit{BlackPropp}$ has a 
negative weight of $-0.50$ which suggests that as the proportion of Black people than typical increase Principal Component 2 will decrease. 
Since Principal Component 2 has a negative weight in our overall model, 
we can say that that there are a higher proportion than average of people who 
are black. Additionally with $\textit{BachelorProp} = 0.46 $ and $\textit{WhiteProp} =0.46$ as they proportions are higher than typical in a block group, Principal Component 2 increases. 

This seems to suggest that eductaion and race are factors that play into spatial mismatch. 


```{r}

Baltimore2$sqrtspatialmismatch<- sqrt(Baltimore2$spatialmismatch)

Baltimore2[is.na(Baltimore2)| Baltimore2 == "Inf"| Baltimore2 == "-Inf"] = NA

M13_PCA_Baltimore<- lm(sqrtspatialmismatch ~ PC2, data= Baltimore2)

summary(M13_PCA_Baltimore)

hist(M13_PCA_Baltimore$resid)
qqnorm(M13_PCA_Baltimore$resid)

car::qqPlot(M13_PCA_Baltimore$residuals)

plot(M13_PCA_Baltimore$resid ~ M13_PCA_Baltimore$fitted)
abline(0,0)

M14_PCA_Baltimore<-lm(spatialmismatch ~ PC2 + I(PC2^2), data = Baltimore2)

```
### Analyze Baltimore Best Model 3

Observing the data, with out best model, we see that our curve doesn't match the data. 
With our low $$ R^{2}_{adj} = 0.01$$ we were not expecting a great model. This does 
suggest that the factors of Race and education ar enot good indicators in prediciting 
spatial mismatch in Baltimore as much as it is in Seattle. 

Moreover,  the cruve line shows that there isn't any pattern to amongt Principal Componet  2 
in regards to Spatial Mismatch 



```{r}

plot(spatialmismatch ~ PC2, data=Baltimore2,
     xlab = "Principal Component 2", 
     ylab = "Spatial Mismatch")
curve((coef(M13_PCA_Baltimore)[1])^2*x^2+2*coef(M13_PCA_Baltimore)[1]*coef(M13_PCA_Baltimore)[2]*x+coef(M13_PCA_Baltimore)[2]^2*x^2, lwd = 2, add =T, col= "blue")
 
ggplot(Baltimore2, aes(x=PC2, y=spatialmismatch)) + geom_point() + geom_smooth()

```

```{r}


#Normalizing Data 

normalized_BaltimoreSpatial<- scale(Baltimore_spatial)

#Compute Correlation Matrix

BaltimoreSpatialcorr_matrix <- cor(normalized_BaltimoreSpatial)
ggcorrplot(BaltimoreSpatialcorr_matrix, colors = c("#d8b365", "#f5f5f5", "#5ab4ac"))

#Application 

PCA_BaltimoreSpatial <-princomp(BaltimoreSpatialcorr_matrix)
summary(PCA_BaltimoreSpatial)

fviz_pca_var(PCA_BaltimoreSpatial, col.var = "black")

PCA_BaltimoreSpatial$loadings[, 1:2]


BaltimoreSpatialpcs<-as.matrix(normalized_BaltimoreSpatial %*% PCA_BaltimoreSpatial$loadings[, 1:2])

BaltimoreSpatialpc <- prcomp(Baltimore_finality3, center = TRUE, scale. = TRUE)

Baltimore2 <- as.data.frame(cbind(spatialmismatch=Baltimore_finality$spatialmismatch, BaltimoreSpatialpc$x[,1:2]))




```



```{r}
# Baltimore Fitting Linear Model PCA


M2_PCA_BaltimoreSpatial <- lm(spatialmismatch ~ PC1 + PC2, data = Baltimore2)

M21_PCA_BaltimoreSpatial <-lm(spatialmismatch ~ PC2, data=Baltimore2)


summary(M2_PCA_BaltimoreSpatial)

summary(M21_PCA_BaltimoreSpatial)



```

Williams Test:


```{r}
library(psych)
r.test(n = 2477, cor(Seattle_finality$spatialmismatch,Seattle_finality3$WhiteProp), cor(Baltimore_finality$spatialmismatch,Baltimore_finality$WhiteProp), n2 = 1939, twotailed =TRUE )
```
###PCA analysis combined

```{r warning=FALSE}

  final<- finality %>%
    select(WhiteProp,AsianProp, HispanicProp, BlackProp, hsdiplomaProp,hsnodiplomaProp,somecollegeProp,college1Prop,assProp,bachelorProp)



#Normalizing Data 

normalized_final<- scale(final)

#Compute Correlation Matrix

corr_matrix <- cor(normalized_final)
ggcorrplot(corr_matrix, colors = c("#d8b365", "#f5f5f5", "#5ab4ac"))

#Application 

PCA_final <-princomp(corr_matrix)


fviz_pca_var(PCA_final, col.var = "cos2", gradient.cols = c("black", "orange", "green"),repel = TRUE)



PCA_final$loadings[, 1:2]

finalpcs<-as.matrix(normalized_final %*% PCA_final$loadings[, 1:2])
finalpc <- prcomp(final, center = TRUE, scale. = TRUE)

final2 <- as.data.frame(cbind(spatialmismatch=finality3$spatialmismatch, finalpc$x[,1:2]))

final3 <- as.data.frame(cbind(spatialmismatch=finality3$spatialmismatch, MSA=finality3$MSA,finalpc$x[,1:2]))


final3<-final3 %>%
  mutate(City =ifelse(MSA=="Seattle",1,0))





```


###Final regression


```{r echo=TRUE}

# Linear Regression with PCA

M3_PCA_final<- lm(spatialmismatch ~ PC1 + PC2, data = final2)

summary(M3_PCA_final)

qqnorm(M3_PCA_final$resid)
hist(M3_PCA_final$resid)


final2$logspatialmismatch <- log(finality3$spatialmismatch)

final2[is.na(final2)| final2 == "Inf"| final2 == "-Inf"] = NA

M31_PCA_final<- lm(logspatialmismatch ~ PC1 + PC2, data = final2)

summary(M31_PCA_final)

car::qqPlot(M31_PCA_final$resid)
hist(M31_PCA_final$resid)










```

### William's Test

```{r eval=FALSE, include=FALSE}
cor(final2$spatialmismatch,final2$PC1)
cor(final2$spatialmismatch,final2$PC2)

r.test(n = 2477, cor(Seattle_finality$spatialmismatch,Seattle_finality$PC1),cor(Baltimore_finality$spatialmismatch,Baltimore_finality$PC1), n2 = 1939, twotailed =TRUE )



```

combine components and then find the correlations. 
spatial mismatch is related to different components and then try cities

combine different cities into one PCA and then run correlations based on PC. 


```{r echo=TRUE}
Variables <- c("WhiteProp", "AsianProp", "HispanicProp", "BlackProp" , "hsdiplomaProp", "hsnodiplomaProp", "somecollegeProp", "College1Prop", "associateProp", "bachelorProp")
Principal_Component_1 <- c( 0.29 , 0.11, -0.37, -0.29, -0.43, -0.26, -0.21, -0.26, -0.10, 0.55 )
Principal_Component_2 <- c(0.57, -0.51, -0.21, -0.33, 0.20, -0.06, 0.25, 0.30, 0.21, -0.11)

Seattle_PCA_Results<-data.frame(Variables,Principal_Component_1, Principal_Component_2)

```


```{r echo=TRUE}

PrincipalComponent_2 <- c(0.43, -0.45, -0.23, -0.36, 0.32, -0.007, 0.07, 0.49, 0.35, -0.14)

Baltimore_PCA_Results<- data.frame(Variables, PrincipalComponent_2)

Baltimore_PCA_Results
```


...


